Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-09-07 Manuscript.Rmd
More work on the results and figure captions

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey & Joseph T. McGuire"
output:
  word_document: 
    reference_docx: Style_reference.docx
  html_document: default
  pdf_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

## tanh-z transformation (variance stabilizing Fisher) and p-values (adjusted and not)
# This takes either a matrix of correlation values (vectors too, but manually compute pvals)
# Normalization approach suggested in network textbook (equation 7.10)
fisherTanh <- function(Data = padjMatrix){

@@

The distribution suggests that both clustering methods were in high agreement across subjects (mean = `r round(mean(RI_all_within$Index), digits = 2)` sd = `r round(sd(RI_all_within$Index), digits = 2)`). The importance of this results lies in that while spectral partitioning forces a bipartition of the network, modularity does not. This adds evidence to the division of these areas into DMN and non-DMN during rest.

Next, we examine how similar the partitions for PCC and mPFC are across subjects. Qualitative inspection of the brain community topologies shows good alignment for PCC, with a topologically similar mPFC pattern that shifts topographically across subjects. The following similarity matrix shows RIs for every pairwise comparison among all subjects for PCC (lower matrix) and PFC (upper) separately. Each row/column element is a single subject.

@@

``` {r Partition comparison: all 2, fig.align="center", fig.width = 4, fig.height = 4, fig.cap = "Figure 4. Similarity matrix showing ARI values among all subjects for PCC (lower triangle) and mPFC (upper triangle) separately. PCC is well aligned across individuals, while mPFC shows clear topographical heterogeneity.", echo=FALSE, message = FALSE, cache = TRUE}
# Across subjects
# Note: the upper and lower triangles can totally be combined before plotting, instead of adding one corrplot over another. Might work on that eventually.
RI_all_between <- comparePartitions(Data = Summaries, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
posteriorIndx <- Summaries[[1]]$y < 0
tempComm_PCC <- lapply(Summaries, "[", i = posteriorIndx, j =)
tempComm_PFC <- lapply(Summaries, "[", i = !posteriorIndx, j =)
RI_PFC <- comparePartitions(Data = tempComm_PFC, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
RI_PCC <- comparePartitions(Data = tempComm_PCC, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
# par(mfrow = c(1,2))
# corrplot(RI_all_between,
#          is.corr = F, 
#          method = "color", 
#          tl.pos = "n", 
#          tl.col = "black", 
#          cl.lim = c(0,1),
#          cl.pos = "b",
#          cl.length = 3)
# mtext("Subjects", side = 2, line = -4.5, cex = 1.5)
# mtext("Subjects", side = 3, line = 4.5, cex = 1.5)
corrplot(RI_PCC,
         is.corr = F, 
         method = "color",
         type = "lower", 
         tl.pos = "n", 
         tl.col = "black", 
         cl.lim = c(0,1),
         cl.length = 3,
         cl.pos = "b")
corrplot(RI_PFC, 
         is.corr = F, 
         method = "color", 
         type = "upper", 
         tl.pos = "n", 
         tl.col = "black",
         cl.pos = "n", 
         add = T)

# Let's compare stuff
testPCC <- RI_PCC[lower.tri(RI_PCC)]
testPFC <- RI_PFC[lower.tri(RI_PFC)]
Overall_PCCvsPFC_permTest <- permute(testPCC, testPFC, statType = mean, paired = T)
Overall_PCCvsPFC_ES <- cohenD(testPCC, testPFC) # caveat: maybe not ideal for bounded values, like RI. BUT VI yields similar results.

The results match the qualitative observations mentioned above, with overall higher agreement for PCC (paired permutation, p = `r Overall_PCCvsPFC_permTest$Pval`; Cohen's D = `r round(Overall_PCCvsPFC_ES, digits = 2)`). However, the heterogeneity of mPFC could arise from noise in the signal. As a way to test for subject specificity, we used each subject's session data to compare day 1 vs day 2 clusters, as well as a session-based analysis.

\bigskip
\bigskip

@@

``` {r Partition comparison: halves, fig.align="center", fig.width = 8, fig.height = 4, fig.cap = "Figure 5. Similarity matrices showing pattern agreement across days for the whole partition (left), and for PCC and mPFC separately (right). These figures show high agreement within (but not between) subjects, with mPFC again showing significant variability.", echo=FALSE, message = FALSE, cache = TRUE}
# Select 10 subjects from the sample to display
selectSubj <- seq(1,20)

@@

# Compare PCC and PFC among all subjects
tempComm_PCC_halves_plot <- lapply(Summaries_halves_plot, "[", i = posteriorIndx, j =)
tempComm_PFC_halves_plot <- lapply(Summaries_halves_plot, "[", i = !posteriorIndx, j =)
RI_PFC_halves <- comparePartitions(Data = tempComm_PFC_halves_plot, Index = "RI", nSubjects = length(Summaries_halves_plot), subjNames = subjList_halves_plot)
RI_PCC_halves <- comparePartitions(Data = tempComm_PCC_halves_plot, Index = "RI", nSubjects = length(Summaries_halves_plot), subjNames = subjList_halves_plot)

# plots
par(mfrow=c(1,2))
corrplot(RI_all_between_halves, 
         is.corr = F, 
         method = "color",
         tl.pos = "n", 
         tl.col = "black",
         tl.cex = 0.8,
         cl.pos = "b",
         cl.length = 3)
corrplot(RI_PCC_halves, 
         is.corr = F,     
         method = "color",
         type = "lower", 
         tl.pos = "n", 
         tl.col = "black", 
         tl.cex = 0.8,
         cl.lim = c(-0.03,1),
         cl.length = 3)
corrplot(RI_PFC_halves, 
         is.corr = F, 
         method = "color",
         type = "upper", 
         tl.pos = "n", 
         tl.col = "black",
         tl.cex = 0.8,
         cl.pos = "n", 
         add = T)

@@

\bigskip
\bigskip

@@

``` {r Partition comparison: session similarity matrices, fig.align="center", echo=FALSE, message = FALSE, cache = TRUE}
# Select 10 subjects from the sample to display
selectSubj <- seq(1,40)

@@

# Across subjects (just for plotting)
RI_all_between_sess <- comparePartitions(Data = Summaries_sess_plot, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
tempComm_PCC_sess <- lapply(Summaries_sess_plot, "[", i = posteriorIndx, j =)
tempComm_PFC_sess <- lapply(Summaries_sess_plot, "[", i = !posteriorIndx, j =)
RI_PFC_sess <- comparePartitions(Data = tempComm_PFC_sess, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
RI_PCC_sess <- comparePartitions(Data = tempComm_PCC_sess, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)

# plot
par(mfrow=c(1,2))
corrplot(RI_all_between_sess, 
         is.corr = F, 
         method = "color", 
         tl.pos = "n", 
         tl.cex = 0.5,
         tl.col = "black", 
         cl.lim = c(-0.03,1),
         cl.pos = "b",
         cl.length = 3)
corrplot(RI_PCC_sess, 
         is.corr = F, 
         method = "color", 
         type = "lower", 
         tl.pos = "n", 
         tl.cex = 0.5,
         tl.col = "black", 
         cl.lim = c(-0.03,1),
         cl.length = 3)
corrplot(RI_PFC_sess, 
         is.corr = F, 
         method = "color", 
         type = "upper", 
         tl.pos = "n", 
         tl.col = "black", 
         cl.pos = "n", 
         add = T)

@@

``` {r Partition comparison: session ratios, fig.align="center", fig.width = 4, fig.height = 4, echo=FALSE, message = FALSE, cache = TRUE}
## Brain storming:
## Within-subject comparisons would be more interesting? Like, comparing the mean RI within-between per subject
## Perhaps a more interresting idea would be to generate an index of variance within/between subject
## These are just attempts at that are not perfect:

@@

ggplot(data = Sess_RIRatios_summary, aes(Region, RI, fill = Region)) + 
          geom_boxplot(show.legend = F) +
          scale_fill_manual(values = Cols[seq(2)]) + 
          ylim(0, max(Sess_RIRatios_summary$RI)) +
          geom_hline(yintercept = 1) +
          labs(y = "RI Ratio", x = "") + 
          theme_classic()
```

@@

``` {r Stability per subject: Overall, fig.align="center", fig.width=8, fig.height=4, echo=FALSE, message = FALSE, cache = TRUE}
# Load files
setwd('./Sliding_window/')
temp <- list.files(pattern = '*Comparisons.csv')
SC <- lapply(temp, read.csv2)
temp <- list.files(pattern = '*Values.csv')
SVals <- lapply(temp, read.csv2)

@@

# Fiedler Vector x proportion of times a vertex was affiliated with DMN
SC_FVvPDMN_plot <- ggplot(data = allSummaries) +
  geom_smooth(aes(FiedlerVec, slidePropDMN, group = SubjID), color = "gray", fill = "gray", show.legend = F, alpha = 0.3) +
  #geom_point(aes(FiedlerVec, slidePropDMN, fill = SubjID), show.legend = F, alpha = 0.05, pch = 21, color = "black") + 
  geom_line(aes(FiedlerVec, FittedVals), alpha = 1, size = 3) +
  labs(x = "Fiedler Vector Values", y = "Proportion of DMN Affiliations") +
  theme_classic()

@@

While some subjects show considerable variability, the overall trend is of fair stability. An interesting feature of spectral partitioning is that the Fiedler vector, which provides signed values to divide a network into positive and negative communities, was tightly related to the proportion of times any given cortical vertex was affiliated with the DMN throughout the fMRI sessions. Meaning, the higher the absolute value in the Fiedler vector for a given vertex, the more persistent its relationship was with its corresponding community in time. The following plot shows this relationship for each subject (gray smoothed lines), as well as the fitted values from a logistic mixed effects model (black line; model: prop. DMN ~ Fiedler vector, subjects as random intercepts; beta = `r round(SC_FVvPDMN_modelSummary$coefficients[2,1], digits = 2)`, SE = `r round(SC_FVvPDMN_modelSummary$coefficients[2,2], digits = 2)`, p < 0.001).

``` {r Stability per subject: Fiedler vector vs prop. DMN plot, fig.align="center", fig.width=4, fig.height=4, echo=FALSE, message = FALSE}
SC_FVvPDMN_plot
```

@@

``` {r Stability per subject: PCC vs mPFC, fig.align="center", fig.width=10, fig.height=4, echo=FALSE, message = FALSE, cache = TRUE}
## Check stability for PCC and mPFC separately
# Get the sliding values for PCC and PFC separately
tempSW_PCC <- lapply(SVals, "[", i = posteriorIndx, j =)
tempSW_PFC <- lapply(SVals, "[", i = !posteriorIndx, j =)

@@

@@

This suggests that, while the agreement of each window with the overall time series was stable for both PCC and mPFC (low SD, right plot), the overall agreement of mPFC is lower (paired permutation, p = `r SW_PCCvPFC_permuteTest$Pval`; Cohen's D = `r round(SW_PCCvPFC_ES, digits = 2)`). This result could be due to signal degeneration in mPFC (that said, it is worth noting that the topological patterns remain mostly the same for both PCC and mPFC throughout the SW analysis). To account for this, we computed the temporal signal to noise ratio (tSNR) for each vertex in each subject's minimally preprocessed fMRI data (i.e. mean of each vertex's time points over their standard deviation). The next plot shows the pooled subject-wise mean tSNR for PCC and mPFC vertices, and indicates that the tSNR is lower for mPFC than PCC, which could negatively affect our results.

``` {r Stability per subject: SNR Comparison, fig.align="center", fig.width=4, fig.height=4, echo=FALSE, message = FALSE, cache = TRUE}
## Putative steps to analyze the SNR of the full time series (PCC vs mPFC)
# Flexibility vs SNR
# Are vertices with lower SNR switching communities more often?
SNR_vs_Flex <- sapply(Summaries, function(data) cor(data$varCoeff, data$slideFlexibility))

@@

# Plot the mPFC SNR vs PCC SNR for each subject (one figure containing all subjects)
snrmeanComp_plot <- ggplot(data = SNR_meanComp, aes(Region, SNR, fill = Region)) +
  geom_boxplot(show.legend = F) +
  scale_fill_manual(values = Cols[seq(2)]) +
  labs(x = "", y = "SNR") +
  theme(text = element_text(size=35)) +
  theme_classic()
snrmeanComp_plot
```

@@

``` {r Stability per subject 4, fig.align="center", fig.width=5, fig.height=4, echo=FALSE, message = FALSE, cache = TRUE}
setwd('./Variance/')
temp <- list.files()
Variance <- lapply(temp, read.csv)

@@

# There is a strong linear relationship between the Fiedler vector and the R-squared values
# These match perfectly on the brain when inspected qualitatively
v1 <- qplot(data = allVariance, FiedlerVec, Rsquared, color = SubjID, fill = SubjID, geom = "smooth", show.legend = F) + labs(x = "Fiedler Vector Values", y = "R-Squared Values") + theme_classic()

@@

``` {r Stability per subject 5, fig.align="center", fig.width=6, fig.height=4, echo=FALSE, message = FALSE}
v4
```

@@

``` {r Stability per subject 6, fig.align="center", fig.width=8, fig.height=4, echo=FALSE, message = FALSE, cache = TRUE}
# grid.arrange(v6, v7, ncol = 2)
par(mfrow = c(1,2))
smoothScatter(allVariance[allVariance$Region == "mPFC",]$Rsquared, allVariance[allVariance$Region == "mPFC",]$tempZ,
              xlab = "R-squared Value",
              ylab = "Z-Coordinate")

smoothScatter(allVariance[allVariance$Region == "mPFC",]$varCoeff, allVariance[allVariance$Region == "mPFC",]$tempZ,
              xlab = "Coefficient of Variation (SNR)",
              ylab = "")
```

@@

``` {r Correlation vs community detection: plot, fig.align="center", echo=FALSE, message = FALSE, cache = TRUE}
## Between halves analysis for correlation and communities (mPFC only)
tempComm_PFC_halves_plot <- lapply(seq_along(tempComm_PFC_halves_plot), function(x) cbind(tempComm_PFC_halves_plot[[x]], Correlation_halves_plot[[x]]))
D1 <- tempComm_PFC_halves_plot
D2 <- lapply(tempComm_PFC_halves_plot, function(data) transform(data, V1 = FiedlerVec)) # the comparison function can only take 1 column name for this stuff
Data <- c(D1,D2)
allHalfCombs <- comparePartitions(Data = Data, MOI = "V1", nSubjects = 40, Index = "Cor", subjNames = c(subjList_halves_plot, subjList_halves_plot))
corrplot(allHalfCombs, 
         title = "Comparing All Community and Correlation mPFC Partitions \n Across All Subjects",
         method = "color", 
         tl.col = "black",
         tl.cex = 0.6,
         cl.pos = "b",
         mar = c(0,0,3,0))
segments(0.5,0.5,0.5,40.5, lwd = 2)
segments(0.5,0.5,40.5,0.5, lwd = 2)
segments(0.5,40.5,40.5,40.5, lwd = 2)
segments(40.5,0.5,40.5,40.5, lwd = 2)
segments(0.5,20.5,40.5,20.5, lwd = 2)
segments(20.5,0.5,20.5,40.5, lwd = 2)

@@

``` {r Correlation vs community detection: comparison plot, fig.align="center", fig.width=4, fig.height=4, echo=FALSE, message = FALSE}
# Day 1 vs Day 2 Spatial Correlation Comparison
prepostComp_plot <- ggplot(data = prepostComp, aes(Approach, Correlation, fill = Approach)) +
  geom_boxplot(show.legend = F) +
  labs(x = "", y = "Correlation Coefficient") +
  ylim(0,1) +
  theme(text = element_text(size=35)) +
  theme_classic()
prepostComp_plot
```

``` {r 16.1.1., fig.align="center", fig.width=6, fig.height=4, echo=FALSE, message = FALSE}

>>>>

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey & Joseph T. McGuire"
output:
  word_document: 
    reference_docx: Style_reference.docx
  pdf_document: default
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

## tanh-z transformation (variance stabilizing Fisher) and p-values (adjusted and not)
# This takes either a matrix of correlation values (vectors too, but manually compute pvals)
# Normalization approach suggested in network textbook (equation 7.10)
fisherTanh <- function(Data = padjMatrix){

@@

``` {r Partition comparison: all 2, echo=FALSE, message = FALSE, cache = TRUE}
# Across subjects
# Note: the upper and lower triangles can totally be combined before plotting, instead of adding one corrplot over another. Might work on that eventually.
RI_all_between <- comparePartitions(Data = Summaries, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
RI_all_between_lowtri <- RI_all_between[lower.tri(RI_all_between)]

posteriorIndx <- Summaries[[1]]$y < 0
tempComm_PCC <- lapply(Summaries, "[", i = posteriorIndx, j =)
tempComm_PFC <- lapply(Summaries, "[", i = !posteriorIndx, j =)
RI_PFC <- comparePartitions(Data = tempComm_PFC, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)
RI_PCC <- comparePartitions(Data = tempComm_PCC, Index = "RI", nSubjects = length(Summaries), subjNames = subjList)

# Let's compare stuff
testPCC <- RI_PCC[lower.tri(RI_PCC)]
testPFC <- RI_PFC[lower.tri(RI_PFC)]
Overall_PCCvsPFC_permTest <- permute(testPCC, testPFC, statType = mean, paired = T)
Overall_PCCvsPFC_ES <- cohenD(testPCC, testPFC) # caveat: maybe not ideal for bounded values, like RI. BUT VI yields similar results.
```

The distribution suggests that both clustering methods were in high agreement across subjects (mean = `r round(mean(RI_all_within$Index), digits = 2)` sd = `r round(sd(RI_all_within$Index), digits = 2)`). The importance of this result lies in that while spectral partitioning forces a bipartition of the network, modularity does not. This adds evidence to the division of these areas into DMN and non-DMN during rest.

Next, we examined how similar the partitions were across subjects by computing the ARI between every pair of subjects, and found that there was considerable agreement overall (mean = `r round(mean(RI_all_between_lowtri), digits = 2)`, SD = `r round(sd(RI_all_between_lowtri), digits = 2)`). However, qualitative inspection of the brain community topologies shows good alignment for PCC, with a topologically similar mPFC pattern that shifts topographically across subjects. To quantify this heterogeneity in mPFC, we performed the same analysis for each region separately (Figure 4). Indeed, this analysis showed that the functional topology of PCC is well maintained across subjects (mean = `r round(mean(testPCC), digits = 2)`, SD = `r round(sd(testPCC), digits = 2)`), while mPFC (mean = `r round(mean(testPFC), digits = 2)`, SD = `r round(sd(testPFC), digits = 2)`) showed significantly lower levels of agreement (paired permutation, p = `r Overall_PCCvsPFC_permTest$Pval`; Cohen's D = `r round(Overall_PCCvsPFC_ES, digits = 2)`).

``` {r Partition comparison: all 2 plot, fig.align="center", fig.width = 4, fig.height = 4, fig.cap = "Figure 4. Similarity matrix showing ARI values among all subjects for PCC (lower triangle) and mPFC (upper triangle) separately. PCC is well aligned across individuals, while mPFC shows clear topographical heterogeneity.", echo=FALSE, message = FALSE, cache = TRUE}

# par(mfrow = c(1,2))
# corrplot(RI_all_between,
#          is.corr = F, 
#          method = "color", 
#          tl.pos = "n", 
#          tl.col = "black", 
#          cl.lim = c(0,1),
#          cl.pos = "b",
#          cl.length = 3)
# mtext("Subjects", side = 2, line = -4.5, cex = 1.5)
# mtext("Subjects", side = 3, line = 4.5, cex = 1.5)
corrplot(RI_PCC,
         is.corr = F, 
         method = "color",
         type = "lower", 
         tl.pos = "n", 
         tl.col = "black", 
         cl.lim = c(0,1),
         cl.length = 3,
         cl.pos = "b")
corrplot(RI_PFC, 
         is.corr = F, 
         method = "color", 
         type = "upper", 
         tl.pos = "n", 
         tl.col = "black",
         cl.pos = "n", 
         add = T)

@@

@@

``` {r Partition comparison: halves, echo=FALSE, message = FALSE, cache = TRUE}
# Select 10 subjects from the sample to display
selectSubj <- seq(1,20)

@@

# Compare PCC and PFC among all subjects
tempComm_PCC_halves_plot <- lapply(Summaries_halves_plot, "[", i = posteriorIndx, j =)
tempComm_PFC_halves_plot <- lapply(Summaries_halves_plot, "[", i = !posteriorIndx, j =)
RI_PFC_halves <- comparePartitions(Data = tempComm_PFC_halves_plot, Index = "RI", nSubjects = length(Summaries_halves_plot), subjNames = subjList_halves_plot)
RI_PCC_halves <- comparePartitions(Data = tempComm_PCC_halves_plot, Index = "RI", nSubjects = length(Summaries_halves_plot), subjNames = subjList_halves_plot)
```

``` {r Partition comparison: halves plot, fig.align="center", fig.width = 6, fig.height = 4, fig.cap = "Figure 5. Similarity matrices showing pattern agreement across days for the whole partition (left), and for PCC and mPFC separately (right). These figures show high agreement within (but not between) subjects, with mPFC again showing significant variability.", echo=FALSE, message = FALSE, cache = TRUE}
# plots
par(mfrow=c(1,2))
corrplot(RI_all_between_halves, 
         is.corr = F, 
         method = "color",
         tl.pos = "n", 
         tl.col = "black",
         tl.cex = 0.8,
         cl.pos = "b",
         cl.length = 3)
corrplot(RI_PCC_halves, 
         is.corr = F,     
         method = "color",
         type = "lower", 
         tl.pos = "n", 
         tl.col = "black", 
         tl.cex = 0.8,
         cl.lim = c(-0.03,1),
         cl.length = 3)
corrplot(RI_PFC_halves, 
         is.corr = F, 
         method = "color",
         type = "upper", 
         tl.pos = "n", 
         tl.col = "black",
         tl.cex = 0.8,
         cl.pos = "n", 
         add = T)

@@

@@

``` {r Partition comparison: session similarity matrices, echo=FALSE, message = FALSE, cache = TRUE}
# Select 10 subjects from the sample to display
selectSubj <- seq(1,40)

@@

# Across subjects (just for plotting)
RI_all_between_sess <- comparePartitions(Data = Summaries_sess_plot, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
tempComm_PCC_sess <- lapply(Summaries_sess_plot, "[", i = posteriorIndx, j =)
tempComm_PFC_sess <- lapply(Summaries_sess_plot, "[", i = !posteriorIndx, j =)
RI_PFC_sess <- comparePartitions(Data = tempComm_PFC_sess, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
RI_PCC_sess <- comparePartitions(Data = tempComm_PCC_sess, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
```

``` {r Partition comparison: session similarity matrices plot, fig.align="center", fig.width = 6, fig.height = 4, fig.cap = "Figure 7. Similarity matrices showing pattern agreement across sessions for the whole partition (left), and for PCC and mPFC separately (right). The pattern from the daily analysis is preserved, with marked individuality for mPFC, but not PCC.", echo=FALSE, message = FALSE, cache = TRUE}
# plot
par(mfrow=c(1,2))
corrplot(RI_all_between_sess, 
         is.corr = F, 
         method = "color", 
         tl.pos = "n", 
         tl.cex = 0.5,
         tl.col = "black", 
         cl.lim = c(-0.03,1),
         cl.pos = "b",
         cl.length = 3)
corrplot(RI_PCC_sess, 
         is.corr = F, 
         method = "color", 
         type = "lower", 
         tl.pos = "n", 
         tl.cex = 0.5,
         tl.col = "black", 
         cl.lim = c(-0.03,1),
         cl.length = 3)
corrplot(RI_PFC_sess, 
         is.corr = F, 
         method = "color", 
         type = "upper", 
         tl.pos = "n", 
         tl.col = "black", 
         cl.pos = "n", 
         add = T)

@@

``` {r Partition comparison: session ratios, fig.align="center", fig.width = 4, fig.height = 4, fig.cap = "Figure 8. Whole group ratio between within- and between-subject ARI for mPFC and PCC based on sessions. Consistent with the daily analysis, both regions show idiosyncrasy, with mPFC being significantly more subject-specific.", echo=FALSE, message = FALSE, cache = TRUE}
## Brain storming:
## Within-subject comparisons would be more interesting? Like, comparing the mean RI within-between per subject
## Perhaps a more interresting idea would be to generate an index of variance within/between subject
## These are just attempts at that are not perfect:

@@

ggplot(data = Sess_RIRatios_summary, aes(Region, RI, fill = Region)) + 
          geom_boxplot(show.legend = F) +
          scale_fill_manual(values = Cols[seq(2)]) + 
          ylim(0, max(Sess_RIRatios_summary$RI)) +
          geom_hline(yintercept = 1) +
          labs(y = "ARI Ratio", x = "") + 
          theme_classic()
```

@@

``` {r Stability per subject: Overall, fig.align="center", fig.width=6, fig.height=4, fig.cap = "Figure 9. ARI between communities computed at each sliding epoch and the original partition for all subjects (left), as well as the distribution of ARI standard deviations per subject (right).", echo=FALSE, message = FALSE, cache = TRUE}
# Load files
setwd('./Sliding_window/')
temp <- list.files(pattern = '*Comparisons.csv')
SC <- lapply(temp, read.csv2)
temp <- list.files(pattern = '*Values.csv')
SVals <- lapply(temp, read.csv2)

@@

@@

While some subjects show considerable variability, the overall trend is of fair stability. An interesting feature of spectral partitioning is that the Fiedler vector was tightly related to the proportion of times any given cortical vertex was affiliated with the DMN throughout the sliding windows. Meaning, the higher the absolute value in the Fiedler vector for a given vertex, the more persistent its relationship was with its corresponding community in time. The following plot shows this relationship for each subject (gray smoothed lines), as well as the fitted values from a logistic mixed effects model (black line; model: prop. DMN ~ Fiedler vector, subjects as random intercepts; beta = `r round(SC_FVvPDMN_modelSummary$coefficients[2,1], digits = 2)`, SE = `r round(SC_FVvPDMN_modelSummary$coefficients[2,2], digits = 2)`, p < 0.001).

``` {r Stability per subject: Fiedler vector vs prop. DMN plot, fig.align="center", fig.width=3.5, fig.height=3.5, fig.cap = "Figure 10. Relationship between the magnitude of Fiedler vector values and the proportion of affiliations of a vertex with DMN throughout 37 sliding windows. Grey lines denote each subject's association, while the black line shows the fit from a mixed effects logistic regression.", echo=FALSE, message = FALSE}
# Fiedler Vector x proportion of times a vertex was affiliated with DMN
SC_FVvPDMN_plot <- ggplot(data = allSummaries) +
  geom_smooth(aes(FiedlerVec, slidePropDMN, group = SubjID), color = "gray", fill = "gray", show.legend = F, alpha = 0.3) +
  #geom_point(aes(FiedlerVec, slidePropDMN, fill = SubjID), show.legend = F, alpha = 0.05, pch = 21, color = "black") + 
  geom_line(aes(FiedlerVec, FittedVals), alpha = 1, size = 3) +
  labs(x = "Fiedler Vector Values", y = "Proportion of DMN Affiliations") +
  theme_classic()

SC_FVvPDMN_plot
```

@@

``` {r Stability per subject: PCC vs mPFC, fig.align="center", fig.width=10, fig.height=4, fig.cap = "Figure 11. ARI between each window and the original pattern for PCC and mPFC separately (left), and the mean per-subject ARI for the whole group (middle). Regardless of the mean ARI, standard deviation remained low for both regions (right).", echo=FALSE, message = FALSE, cache = TRUE}
## Check stability for PCC and mPFC separately
# Get the sliding values for PCC and PFC separately
tempSW_PCC <- lapply(SVals, "[", i = posteriorIndx, j =)
tempSW_PFC <- lapply(SVals, "[", i = !posteriorIndx, j =)

@@

This suggests that, while the agreement of each window with the overall time series was stable for both PCC and mPFC (low SD, right plot), the overall agreement of mPFC is lower (paired permutation, p = `r SW_PCCvPFC_permuteTest$Pval`; Cohen's D = `r round(SW_PCCvPFC_ES, digits = 2)`).

@@

### *Potential Noise Effects*

This result could be due to signal degeneration in mPFC (that said, it is worth noting that the topological patterns remain mostly the same for both PCC and mPFC throughout the SW analysis). To account for this, we computed the temporal signal to noise ratio (tSNR) for each vertex in each subject's minimally preprocessed fMRI data (i.e. mean of each vertex's time points over their standard deviation). The next plot shows the pooled subject-wise mean tSNR for PCC and mPFC vertices, and indicates that the tSNR is lower for mPFC than PCC, which could negatively affect our results.

``` {r Noise Effects: tSNR Comparison, fig.align="center", fig.width=3.5, fig.height=3.5, fig.cap = "Figure 12. Mean tSNR for mPFC and PCC across participants. This measure was derived from the minimally preprocessed fMRI data from the HCP, and thus shows the base level of reltive noise in each region.", echo=FALSE, message = FALSE, cache = TRUE}
## Putative steps to analyze the SNR of the full time series (PCC vs mPFC)
# Flexibility vs SNR
# Are vertices with lower SNR switching communities more often?
SNR_vs_Flex <- sapply(Summaries, function(data) cor(data$varCoeff, data$slideFlexibility))

@@

# Plot the mPFC SNR vs PCC SNR for each subject (one figure containing all subjects)
snrmeanComp_plot <- ggplot(data = SNR_meanComp, aes(Region, SNR, fill = Region)) +
  geom_boxplot(show.legend = F) +
  scale_fill_manual(values = Cols[seq(2)]) +
  labs(x = "", y = "tSNR") +
  theme(text = element_text(size=35)) +
  theme_classic()
snrmeanComp_plot
```

@@

``` {r Noise Effects: Per-subject Fiedler by R-sqd, fig.align="center", fig.width=4, fig.height=4, fig.cap = "Figure 13. Fiedler vector values are linearly related to the total variance explained by each vertex with respect to the mean DMN time series.", echo=FALSE, message = FALSE, cache = TRUE}
setwd('./Variance/')
temp <- list.files()
Variance <- lapply(temp, read.csv)

@@

# There is a strong linear relationship between the Fiedler vector and the R-squared values
# These match perfectly on the brain when inspected qualitatively
v1 <- ggplot(data = allVariance) + 
  geom_smooth(aes(FiedlerVec, Rsquared, group = SubjID), fill = "gray70", color = "gray30", show.legend = F) + 
  labs(x = "Fiedler Vector Values", y = "R-Squared Values") + 
  theme_classic()

@@

``` {r Noise Effects: Mean R-sqd mPFC and PCC, fig.align="center", fig.width=6, fig.height=4, fig.cap = "Figure 14. Differences in mean R-squared values for each region per participant. The figure shows that PCC vertices tend to be more coupled with the overall dynamics of DMN across subjects.", echo=FALSE, message = FALSE}
v4
```

@@

``` {r Noise Effects: R-sqd vs tSNR, fig.align="center", fig.width=6, fig.height=4, fig.cap = "Figure 15. Vertex tSNR values across all subjects are correlated with their position along the z-axis in mPFC, with vmPFC showing the least tSNR (left). However, R-square values do not show this strong relationship (right), suggesting that the coupling of each vertex is not necessarily influenced by signal noise.", echo=FALSE, message = FALSE, cache = TRUE}
# grid.arrange(v6, v7, ncol = 2)
# par(mfrow = c(1,2))
# smoothScatter(allVariance[allVariance$Region == "mPFC",]$Rsquared, allVariance[allVariance$Region == "mPFC",]$tempZ,
#               xlab = "R-squared Value",
#               ylab = "Z-Coordinate")
# 
# smoothScatter(allVariance[allVariance$Region == "mPFC",]$varCoeff, allVariance[allVariance$Region == "mPFC",]$tempZ,
#               xlab = "Coefficient of Variation (SNR)",
#               ylab = "")

tempVariance <- allVariance[allVariance$Region == "mPFC", ]
s1 <- ggplot(data = allVariance, aes(Rsquared, tempZ)) +
  geom_bin2d(bins = 250) +
  scale_fill_continuous(low = "grey30", high = "grey70") +
  geom_density2d(color = Cols[2]) +
  labs(x = "R-squared value", y = "Z-Coordinate") +
  theme_classic()

s2 <- ggplot(data = allVariance, aes(varCoeff, tempZ)) +
  geom_bin2d(bins = 250) +
  scale_fill_continuous(low = "grey30", high = "grey70") +
  geom_density2d(color = Cols[2]) +
  labs(x = "tSNR", y = "Z-Coordinate") +
  theme_classic()

grid.arrange(s2, s1, ncol = 2)
```

@@

``` {r Correlation vs community detection: compute, fig.align="center", echo=FALSE, message = FALSE, cache = TRUE}
## Between halves analysis for correlation and communities (mPFC only)
tempComm_PFC_halves_plot <- lapply(seq_along(tempComm_PFC_halves_plot), function(x) cbind(tempComm_PFC_halves_plot[[x]], Correlation_halves_plot[[x]]))
D1 <- tempComm_PFC_halves_plot
D2 <- lapply(tempComm_PFC_halves_plot, function(data) transform(data, V1 = FiedlerVec)) # the comparison function can only take 1 column name for this stuff
Data <- c(D1,D2)
allHalfCombs <- comparePartitions(Data = Data, MOI = "V1", nSubjects = 40, Index = "Cor", subjNames = c(subjList_halves_plot, subjList_halves_plot))
```

``` {r Correlation vs community detection: plot, fig.align="center", fig.width = 4, fig.height = 4, fig.cap = "Figure 16. Correlation matrix comparing the inter-day spatial stability of maps derived from seed-based functional connectivity (7m-based) and the Fiedler vector. The top-left quadrant represents seed-based FC maps, and the bottom-right the Fiedler vector.", echo=FALSE, message = FALSE, cache = TRUE}
corrplot(allHalfCombs, 
         #title = "Comparing All Community and Correlation mPFC Partitions \n Across All Subjects",
         method = "color", 
         tl.col = "black",
         tl.cex = 0.6,
         cl.pos = "b",
         mar = c(0,0,3,0))
segments(0.5,0.5,0.5,40.5, lwd = 2)
segments(0.5,0.5,40.5,0.5, lwd = 2)
segments(0.5,40.5,40.5,40.5, lwd = 2)
segments(40.5,0.5,40.5,40.5, lwd = 2)
segments(0.5,20.5,40.5,20.5, lwd = 2)
segments(20.5,0.5,20.5,40.5, lwd = 2)

@@

``` {r Correlation vs community detection: comparison plot, fig.align="center", fig.width=4, fig.height=4, fig.cap = "Figure 17. Day 1 vs Day 2 oefficients of spatial correlations for every method, as well as between methods. Community detection through spectral partitioning provides the most stable inter-day estimates, even though both methods show good level of agreement.", echo=FALSE, message = FALSE}
# Day 1 vs Day 2 Spatial Correlation Comparison
prepostComp_plot <- ggplot(data = prepostComp, aes(Approach, Correlation, fill = Approach)) +
  geom_boxplot(show.legend = F) +
  labs(x = "", y = "Correlation Coefficient") +
  ylim(0,1) +
  theme(text = element_text(size=35)) +
  theme_classic()
prepostComp_plot
```
